---
title: "Lexicon home"
author: "Luisa M. Mimmi"
date: "Last run: `r format(Sys.time(), '%F')`"
output: 
  html_document:
    code_folding: hide
editor_options: 
  chunk_output_type: console
bibliography: buzzwords.bib
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)

knit_print.data.frame <- function(x, ...) {
  res <- paste(c('', '', kable_styling(kable(x, booktabs = TRUE))), collapse = '\n')
  asis_output(res)
}

registerS3method("knit_print", "data.frame", knit_print.data.frame)
registerS3method("knit_print", "grouped_df", knit_print.data.frame)

knitr::opts_chunk$set(fig.retina = 3,
                      tidy.opts = list(width.cutoff = 120),  # For code
                      options(width = 90),  # For output
                      fig.asp = 0.618, fig.width = 7, 
                      fig.align = "center", out.width = "85%")

options(dplyr.summarise.inform = FALSE,
        knitr.kable.NA = "")
```

```{r load-libraries-data, message=FALSE, warning=FALSE}
# Load required packages 
if (!require("pacman")) install.packages('pacman' )
library(pacman) # for loading packages
p_load(here, tidyverse, tibble, janitor,
       emo, readr, 
       gt,
       #captioner,
       knitr, 
       flextable, # better bc OK in html & word
       officer, # to save flextable in word
       shiny, 
       stringi,
       wordcountaddin # You may have to download this
       )  
p_load_gh("crsh/citr") # not on CRAN (as of Nov 2021!)
#p_load_gh("benmarwick/wordcountaddin")

# library(tidyverse)
# library(targets)
# library(yardstick)
# library(broom)
# library(patchwork)
# library(scales)
# library(here)

# 

# # Load data
# # Need to use this withr thing because tar_read() and tar_load() need to see the
# # _targets folder in the current directory, but this .Rmd file is in a subfolder
# withr::with_dir(here::here(), {
#   source(tar_read(plot_funs))
#   
#   canary_testing_lagged <- tar_read(panel_testing_lagged)
#   
#   # Load big list of models
#   model_df <- tar_read(model_df) %>% 
#     filter(str_detect(model, "baseline") | str_detect(model, "v2csreprss"))
#   
#   # Load actual model objects
#   tar_load(c(m_pts_baseline_train, m_pts_v2csreprss_train,
#              m_pts_total_train, m_pts_advocacy_train, 
#              m_pts_entry_train, m_pts_funding_train,
#              m_lhr_baseline_train, m_lhr_v2csreprss_train,
#              m_lhr_total_train, m_lhr_advocacy_train, 
#              m_lhr_entry_train, m_lhr_funding_train))
# })
# 
# pts_levels <- levels(canary_testing_lagged$PTS_factor)
# 
# # Returns a data frame of predicted probabilities with actual = 1 
# # when the response outcome happens
# match_actual <- function(x, pred, actual) {
#   pred %>% 
#     select(fitted = {{x}}) %>% 
#     mutate(actual = as.numeric(actual == colnames(pred[x])),
#            plot_level = colnames(pred[x]))
# }
```

# At the core of reduction

I was really struck the first time I realized how the Latin motto I had heard since childhood: "Mens sana in corpore sano", was actually a partial quotation used in 1861 by the Englishman John Hulley, as a motto for his Liverpool Athletic Club. In fact, it was extrapolated from a longer sentence by Juvenal poet: "Orandum est ut sit mens sana in corpore sano" (Satire X, 356). Taken in its entirety, the statement has quite a different meaning than its widely used shortened version: for one thing, it refers prayer and not to sport.

> In alcuni casi la deviazione del significato si evidenzia mediante la soppressione pura e semplice del termine (1984): in tali casi dunque l'analisi non insiste sull'aspetto riduzionista. Le deviazioni, laddove presenti, sono dettate soprattutto da intenti ideologici, ossia dalla sovrapposizione di una griglia interpretativa con maglie piuttosto rigide sui dati oggettivi della realtà. L'effetto della sovrapposizione di questa griglia interpretativa chiamata ideologia risulta in tal modo una riduzione del campo percettivo ed interpretativo della realtà. Per questa ragione le deviazioni semantiche sono state denominate 'riduzionismi'. [@garbini_uscire_2003]

# Some important definitions

## Hermeneutics

Hermeneutics: from Greek *ἑρμηνεία, ας, ἡ* (hermēneia), deriving from the verb *ἑρμηνεύω* (hermēneuō) is the science and practice of interpretation. Over history, it has been mainly applied to sacred or juridical texts. Two are the main approaches:
  1.  reconstructing the original intention of the authors
  2.  adapting the interpretation based on the person who receives the text

## Adage

An *adage* is an ancient saying or maxim, brief and sometimes mysterious, that has become accepted as conventional wisdom. In classical rhetoric, an adage is also known as a *rhetorical proverb* or *paroemia*. Often it's a type of metaphor. It can express the values of a culture. [@nordquist_words_2018]

> **Example(s)**: "The early bird gets the worm", "Better late than never."

## Slogan

The English word *slogan* has a Scottish Gaelic origin and derives from the combination of *sluagh* (army) + *gairm* (shout), i.e. a "battle cry". Nowadays, it signifies a short, memorable and concise phrase used for marketing or political campaigns. Marketing slogans are often called *taglines* in the United States

> **Example(s)**: "...", "..."

# Reference
