---
title: "Process and merge data"
author: "Luisa M. Mimmi"
date: "Last run: `r format(Sys.time(), '%F')`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_depth: 3
editor_options: 
  chunk_output_type: console
bibliography: slogan.bib
---

<i class="fa fa-refresh" style="color: firebrick"></i> Work in progress

# Intro

### --- Some text analytics definitions

-   **Corpus** \~ a collection of similar documents \| objects that typically contain raw strings annotated with additional metadata and details
-   **Document-term matrix** \~ sparse matrix describing a collection (i.e., a corpus) of documents with one row for each document and one column for each term. (`WDR_com`)
-   **Document** \~ collection of sentences
-   **Token** \~ is a meaningful unit of text, such as a *word*, that we are interested in using for analysis
-   **Tokenization** \~ splitting text into tokens
-   **Bigrams**/**n-gram** \~ sentence, or paragraph
-   **Tidy text** \~ "a table with one-token-per-row"
-  **Stemming** \~ A _stem_ is the base part of a word to which affixes can be attached for derivatives. 


### --- Some commonly used NLP scenarios

-   **Text Classification** \~ a supervised learning method of learning and predicting the category or the class of a document given its text content.
-   **Named Entity Recognition** \~ NER is the task of classifying words or key phrases of a text into predefined entities of interest.
-   **Text Summarization** \~ a language generation task of summarizing the input text into a shorter paragraph of text. 
-   **Entailment** \~ the task of classifying the binary relation between two natural-language texts, text and hypothesis, to determine if the text agrees with the hypothesis or not.
-   **Question Answering** \~ QA is the task of retrieving or generating a valid answer for a given query in natural language, provided with a passage related to the query.
-   **Sentence Similarity** \~ the process of computing a similarity score given a pair of text documents.
-   **Embeddings** \~ the process of converting a word or a piece of text to a continuous vector space of real number, usually, in low dimension.
-   **Sentiment Analysis** \~ Provides an example of train and use Aspect Based Sentiment Analysis with Azure ML and Intel NLP Architect .


### --- Text processing in R

-   {tidytext}
-   {tm}  <!-- tm_map(corpus, FUN, ...) -->
-   {SnowballC}  <!--4 stemming  -->
-   {OpenNLP}
-   {RWeka}
-   {languageR}
-   {korPus}



```{r setup_inp, echo=FALSE}
# Pckgs -------------------------------------
if (!require ("pacman")) (install.packages("pacman"))
#p_install_gh("luisDVA/annotater")
#p_install_gh("HumanitiesDataAnalysis/hathidy")
# devtools::install_github("HumanitiesDataAnalysis/HumanitiesDataAnalysis") 
pacman::p_load(here,
       annotater,
       tidyverse, 
       skimr,
       scales,
       httr, 
       citr,
       paint, 
       DT # an R interface to the JavaScript library DataTables
       # jsonlite, 
       # XML, 
       # xml2, 
       # oai, # R client to work with OAI-PMH 
       )

pacman::p_load(
  tm, # Text Mining Package
  tidytext, # Text Mining using 'dplyr', 'ggplot2', and Other Tidy Tools
  sjmisc, # Data and Variable Transformation Functions
  ggraph, # An Implementation of Grammar of Graphics for Graphs and Networks
  widyr, # Widen, Process, then Re-Tidy Data
  SnowballC, # Snowball Stemmers Based on the C 'libstemmer' UTF-8 Library
  HumanitiesDataAnalysis, # Data and Code for Teaching Humanities Data Analysis
  sentencepiece # Text Tokenization using Byte Pair Encoding and Unigram Modelling
)

# Function(s) -------------------------------------

# Data -------------------------------------
 
```


#### -- Set stopwords [more...]

```{r stop}
# --- set up stop words
stop_words <- as_tibble(stop_words) %>% # in the tidytext dataset 
  add_row(word = "WDR", lexicon = NA_character_) %>%
  add_row(word = "world", lexicon = NA_character_) %>%
  add_row(word = "report", lexicon = NA_character_) %>%
  add_row(word = "development", lexicon = NA_character_) %>%
  add_row(word = "1978", lexicon = NA_character_) %>%
  add_row(word = "1979", lexicon = NA_character_) %>%
  add_row(word = "1980", lexicon = NA_character_) %>%
  add_row(word = "1981", lexicon = NA_character_) %>%
  add_row(word = "1982", lexicon = NA_character_) %>%
  add_row(word = "1983", lexicon = NA_character_) %>%
  add_row(word = "1984", lexicon = NA_character_) %>%
  add_row(word = "1985", lexicon = NA_character_) %>%
  add_row(word = "1986", lexicon = NA_character_) %>%
  add_row(word = "1987", lexicon = NA_character_) %>%
  add_row(word = "1988", lexicon = NA_character_) %>%
  add_row(word = "1989", lexicon = NA_character_) %>%
  add_row(word = "1990", lexicon = NA_character_) %>%
  add_row(word = "1991", lexicon = NA_character_) %>%
  add_row(word = "1992", lexicon = NA_character_) %>%
  add_row(word = "1993", lexicon = NA_character_) %>%
  add_row(word = "1994", lexicon = NA_character_) %>%
  add_row(word = "1995", lexicon = NA_character_) %>%
  add_row(word = "1996", lexicon = NA_character_) %>%
  add_row(word = "1997", lexicon = NA_character_) %>%
  add_row(word = "1998", lexicon = NA_character_) %>%
  add_row(word = "1999", lexicon = NA_character_) %>%
  add_row(word = "2000", lexicon = NA_character_) %>%
  add_row(word = "2001", lexicon = NA_character_) %>%
  add_row(word = "2002", lexicon = NA_character_) %>%
  add_row(word = "2003", lexicon = NA_character_) %>%
  add_row(word = "2004", lexicon = NA_character_) %>%
  add_row(word = "2005", lexicon = NA_character_) %>%
  add_row(word = "2006", lexicon = NA_character_) %>%
  add_row(word = "2007", lexicon = NA_character_) %>%
  add_row(word = "2008", lexicon = NA_character_) %>%
  add_row(word = "2009", lexicon = NA_character_) %>%
  add_row(word = "2010", lexicon = NA_character_) %>%
  add_row(word = "2011", lexicon = NA_character_) %>%
  add_row(word = "2012", lexicon = NA_character_) %>%
  add_row(word = "2013", lexicon = NA_character_) %>%
  add_row(word = "2014", lexicon = NA_character_) %>%
  add_row(word = "2015", lexicon = NA_character_) %>%
  add_row(word = "2016", lexicon = NA_character_) %>%
  add_row(word = "2017", lexicon = NA_character_) %>%
  add_row(word = "2018", lexicon = NA_character_) %>%
  add_row(word = "2019", lexicon = NA_character_) %>%
  add_row(word = "2020", lexicon = NA_character_) %>%
  add_row(word = "2021", lexicon = NA_character_) %>%
  add_row(word = "2022", lexicon = NA_character_)
```

# A. World Development Reports (WRDs)

+ DATA https://datacatalog.worldbank.org/search/dataset/0037800 
+ INSTRUCTIONS https://documents.worldbank.org/en/publication/documents-reports/api
+ Following: [@kaye_ella_2019; @robinson_words_2017; @robinson_1_2022]

## A1. Data loading & cleaning
Ingestion of WDR data/metadata is done in `./_my_stuff/WDR-data-ingestion.Rmd` and the result saved as `WDR.rds` <-- (Being somewhat computational intensive, I only did it once.) 

Ingestion of WDR lists of subjects was obtained in .xlsx, but presents issues...
```{r load}
# WRD metadata taken with API get (issues) 
WDR <- readr::read_rds(here::here("data", "raw_data", "WDR.rds" )) %>% 
  # Extract only the portion of string AFTER the backslash {/}
  mutate(id = as.numeric(stringr::str_extract(doc_mt_identifier_1, "[^/]+$"))) %>% 
  dplyr::relocate(id, .before = doc_mt_identifier_1) %>% 
  mutate(url_keys = paste0("https://openknowledge.worldbank.org/handle/10986/", id , "?show=full"))  %>% 
 # eliminate NON WDR book
  dplyr::filter(id != "2586") 

# WRD subject/date_issued taken by manual review 
WDR_subjects <- readxl::read_excel(here::here("data", "raw_data", 
                                              "WDR_subjects_corrected2010_2011.xlsx")) %>%
  drop_na(id) %>% 
 # eliminate NON WDR book
  dplyr::filter(id != "2586") 

# delete empty cols 
ColNums_NotAllMissing <- function(df){ # helper function
  as.vector(which(colSums(is.na(df)) != nrow(df)))
}

WDR_subjects <- WDR_subjects  %>% 
  select(ColNums_NotAllMissing(.))
 # # convert all columns that start with "subj_" to lowercase
 # WDR_subjects[3:218] <- sapply(WDR_subjects[3:218], function(x) tolower(x))

# join
WDR_com <- left_join(WDR, WDR_subjects, by = "id") %>% 
  dplyr::relocate(date_issued, .before = id ) %>% 
  # drop useles clmns 
  dplyr::select(#-doc_mt_identifier_1, 
                -doc_mt_identifier_2, -doc_mt_date, 
                -doc_mt_subject, -doc_mt_creator, -doc_mt_set_spec) %>% 
  # dplyr::relocate(url_keys, .after = subj_216 ) %>% 
  dplyr::rename(abstract = doc_mt_description) %>% 
  # correct titles -> portion after {:}
  dplyr::mutate(., title = str_extract(doc_mt_title,"[^:]+$")) %>% 
  dplyr::relocate(title, .after = id)  %>% 
  dplyr::rename(title_miss = doc_mt_title) %>% 
  dplyr::mutate(title_miss = case_when(
    str_starts(title, "World Development Report") ~ "Y",
    TRUE ~ NA_character_) 
  ) %>% 
  dplyr::mutate(subject_miss = if_else(is.na(subj_1), 
                                       "Y", 
                                       NA_character_)) %>% 
  dplyr::relocate(subject_miss, .after = title_miss)   

# convert all columns that start with "subj_" to lowercase (maybe redundant)
WDR_com[, grep("^subj_", names(WDR_com))] <- sapply(WDR_com[, grep("^subj_", names(WDR_com))], function(x) tolower(x))

# combine all `subj_...` vars into a vector separated by comma
col_subj <- names(WDR_com[, grep("^subj_", names(WDR_com))] )

WDR_com <- WDR_com %>% tidyr::unite(
  col = "subj_all", 
  subj_1:subj_46, 
  sep = ",",
  remove = FALSE) %>% 
  arrange(date_issued)
paint(WDR_com)

```

#### -- Some manual correction of wrong metadata 
```{r meta_correct }
# adding actual titles 
#WDR_com[WDR_com$date_issued == "1978", c("date_issued", "id", "title")]
WDR_com$title[WDR_com$id == "5961"] <- "Prospects for Growth and Alleviation of Poverty"

#WDR_com[WDR_com$date_issued == "1979", c("date_issued", "id", "title")]
WDR_com$title[WDR_com$id == "5962"] <- "Structural Change and Development Policy"

#WDR_com[WDR_com$date_issued == "1980", c("date_issued", "id", "title")]
WDR_com$title[WDR_com$id == "5963"] <- "Poverty and Human Development"

#WDR_com[WDR_com$date_issued == "1981", c("date_issued", "id", "title")]
WDR_com$title[WDR_com$id == "5964"] <- "National and International Adjustment"

#WDR_com[WDR_com$date_issued == "1982", c("date_issued", "id", "title")]
WDR_com$title[WDR_com$id == "5965"] <- "Agriculture and Economic Development"

#WDR_com[WDR_com$date_issued == "1983", c("date_issued", "id", "title")]
WDR_com$title[WDR_com$id == "5966"] <- "Management in Development"

#WDR_com[WDR_com$date_issued == "1984", c("date_issued", "id", "title")]
WDR_com$title[WDR_com$id == "5967"] <- "Population Change and Development"

#WDR_com[WDR_com$date_issued == "1985", c("date_issued", "id", "title")]
WDR_com$title[WDR_com$id == "5968"] <- "International Capital and Economic Development"

#WDR_com[WDR_com$date_issued == "1986", c("date_issued", "id", "title")]
WDR_com$title[WDR_com$id == "5969"] <- "Trade and Pricing Policies in World Agriculture"

#WDR_com[WDR_com$date_issued == "1987", c("date_issued", "id", "title")]
WDR_com$title[WDR_com$id == "5970"] <- "Industrialization and Foreign Trade"

#WDR_com[WDR_com$date_issued == "1988", c("date_issued", "id", "title")]
WDR_com$title[WDR_com$id == "5971"] <- "Public Finance in Development"

# wrong year 
#WDR_com[WDR_com$date_issued %in% c( "2011","2012","2013", "2014", "2015"), c("date_issued", "id", "title")]
WDR_com$date_issued[WDR_com$id == "11843"] <- "2013"
WDR_com$date_issued[WDR_com$id == "16092"] <- "2014"

```



## A1.a Troubleshooting some obs

> PROBLEM: some of the subjects collections are evidently wrong (either they are the same of another WDR or the list is impossibly long)

> MY SOLUTION #1: I took them manually from the website "elibrary"  https://elibrary.worldbank.org/action/showPublications?SeriesKey=b02  

But, there still is WDR 2011 ("Conflict, Security, and Development") which misses keywords 

> MY SOLUTION #2: I take the abstract and I create my own "plausible list" of subjects

### --- Extrapolate subjects from abstracts - for record with missing subjects/keywords

(\*) There will remain a problem: this corrected records have tokens and not bi\|n-grams (which make more sense)!

```{r wrongWDRs}
# --- identify wrong subject obs
WDR_wr <- WDR_com %>% 
  filter( subject_miss == "Y") 
# names(WDR_com)
```

#### -- WDR caseid = 4389

```{r fix4389}
# --- tokenize abstract 
WDR_4389 <- WDR_wr %>% 
  filter(id =="4389") %>% 
  select (abstract) %>% as_tibble() %>% 
  tidytext::unnest_tokens(word, abstract) # -> 251 words 

# --- remove stop words
# --- isolate meaningful tokens
WDR_4389 <- WDR_4389 %>% 
  anti_join( stop_words , by = "word") %>%   # -> 131  words 
  # group and count 
  count(word, sort = T)  # -> 101  words 

# rename column in result corrected   
WDR_4389_w <- t(WDR_4389) %>% as_tibble()  
names(WDR_4389_w) <- gsub(x = names(WDR_4389_w), pattern = "\\V", replacement = "subj_")
 # --- graph words
  # p <- WDR_4389 %>% count(word, sort = TRUE) %>%
  #   filter(n > 600) %>%
  #   mutate(word = reorder(word, n)) %>%
  #   ggplot(aes(n, word)) +
  #   geom_col() +
  #   labs(y = NULL)
  # p  
  #   

# --- replace as subjects 
WDR_4389_w  <- WDR_4389_w %>%  
  filter(row_number() == 1 ) %>% 
  mutate(id = 4389) %>% 
  relocate(id, .before = subj_1)

# names(WDR_4389_w)
WDR_4389_w <- WDR_4389_w %>% 
  tidyr::unite(
    col = "subj_all", 
    subj_1:subj_100, 
    sep = ",",
    remove = FALSE) 


#names(WDR_com)
WDR_com_2 <- WDR_com %>% 
  relocate(subj_1 ,.after = subj_all)

# # create a vector of column names to add to reach the n length(WDR_com)
# col_names <-  as.vector(paste0('subj_', length(WDR_4389_w):105))
# 
# # make a df with those cols and 1 row made of Nas values 
# to_add <- bind_rows(setNames(rep("", length(col_names)), col_names))[NA_character_, ]
# 
# # --- pad until subj_216 with nas.........  
# WDR_4389_w_pad <- bind_cols(WDR_4389_w, to_add)

#### -- NOW: replace corrected single WDR case into master df 
  # # initial check ---
  # WDR_com$subj_1[WDR_com$id == 4389] 
  # WDR_com$subj_3[WDR_com$id == 4389] 
  # #--- 

#id <-  4389 
col_subj_names <- names(WDR_4389_w)[-(1)] # without "id"  "subj_all"

# pick the id 
i <- 4389

# # --- function --- NO JOY!
# for (j in 1:length(col_subj_names)) {
#   col <-  col_subj_names[j]
#   # print(col) # nolint
#   WDR_com %>%
#     dplyr::filter (id == i) %>% 
#     dplyr::mutate (., col = WDR_4389_w_pad$col) 
#   WDR_com_2 <- WDR_com
# }

# ---------# Solution from SO guy 
# r is vectorized so 
WDR_com_2[WDR_com_2$id %in% WDR_4389_w$id, 9:55] <- WDR_4389_w[, 2:48] 


```

\* I cut some ...so this record WDR_4389 will be incomplete 

<!-- #### -- WDR caseid = 2586 -->
 
#### -- save wdr and cleanenv
```{r cleanenv}
wdr <- WDR_com_2 %>% 
  select(-title_miss)

dataDir <- fs::path_abs(here::here("data","derived_data"))
fileName <- "/wdr.rds"
Dir_File <- paste0(dataDir, fileName)
write_rds(x = wdr, file = Dir_File)

# ls objects
list_old_WDR <-  ls(# pattern = "^WDR", 
                    all.names = TRUE)
list_old_WDR
rm(list = list_old_WDR)
```

## A2 Tokenization (of WDRs' abstract)

```{r}
wdr <- readr::read_rds(here::here("data", "derived_data", "wdr.rds" ))
```

#### -------------- review | ------------------

### --- Starting from 1 abstract

```{r}
abs_1 <- as_tibble(wdr$abstract[1] )
```

Break the text [e.g. 1 abstract] into individual tokens

```{r}
# --- modo 1 
# abs_1_t2 <- abs_1 %>%
#   stringr::str_split("[^A-Za-z]")  # gives a list 

# --- modo 2
#abs_1_t <-  abs_1 %>% 
#  # unnest (newColName, oldColName)
#  unnest_tokens(word, value ) # --> tidytext

```

### --- The choices of tokenization

    + Should words be lowercased?
    + Should punctuation be removed?
    + Should numbers be replaced by some placeholder?
    + Should words be stemmed (also called lemmatization).
    + Should bigrams or other multi-word phrase be used instead of or in addition to single word phrases?
    + Should stopwords (the most common words) be removed?
    + Should rare words be removed?

```{r}
# features I can add to `unnest_tokens`
tok_feat_l <- list(
  # all 2 lowercase 
  abs_1 %>% unnest_tokens(word, value) %>% select(lowercase = word),
  # This function`SnowballC::wordStem` extracts the stems of each of the given words in the vector.
  abs_1 %>% unnest_tokens(word, value) %>% rowwise() %>% mutate(word = SnowballC::wordStem(word)) %>% select(stemmed = word),
  # keep uppercase if there are 
  abs_1 %>% unnest_tokens(word, value, to_lower = F) %>% select(uppercase = word),
  # keep punctuation {default is rid} 
  abs_1 %>% unnest_tokens(word, value, to_lower = F, strip_punc = FALSE) %>% select(punctuations = word),
  # bigram
  abs_1 %>% unnest_tokens(word, value, token = "ngrams", n = 2, to_lower = F) %>% select(bigrams = word)
) 

# Return a data frame created by column-binding.
tok_feat_df <- map_dfc(tok_feat_l  , ~ .x %>% head(10))
tok_feat_df

# my choice 
abs_1_t_mod <- abs_1 %>% 
  # no punctuation, yes capitalized
  unnest_tokens(word, value, to_lower = F, strip_punc = TRUE) %>% # 249 obs
  # exclude stopwords 
  anti_join(stop_words) # 109 obs

head(abs_1_t_mod, 15)
```

### --- Tokenizing ALL abstracts

```{r}
# isolate only abstracts 
abs_all <- wdr %>% 
  dplyr::select(id, date_issued, title, abstract)

abs_all_token <- abs_all %>% 
  unnest_tokens(output =  word,
                input = abstract,
                to_lower = T, # otherwise cannot match the stop_words  
                strip_punc = TRUE) %>% #10018
  anti_join(stop_words) # 4786
```

### ---  Plotting obtained tokens (all abstractS) ----------------

```{r}
## Wordcounts/plots --------------------------------------------------------

# Count words
wordcounts <- abs_all_token %>%
  group_by(word) %>%
  summarize(n = n()) %>%
  arrange(-n)

#wordcounts %>% head(5)
# plot
wordcounts <- wordcounts %>%
  mutate(rank = rank(-n)) %>%
  filter(n > 2, word != "")

ggplot(wordcounts) + 
  aes(x = rank, y = n, label = word) + 
  geom_text() # ugly!!!

# Putting logarithmic scales on both axes reveals something interesting about the way that data is structured; this turns into a straight line.
ggplot(wordcounts) +
  aes(x = rank, y = n, label = word) +
  geom_point(alpha = .3, color = "grey") +
  geom_text(check_overlap = TRUE) +
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  labs(title = "Zipf's Law",
       subtitle="The log-frequency of a term is inversely correlated with the logarithm of its rank.")
# ...the logarithm of rank decreases linearly with the logarithm of count
# This is “Zipf’s law:” the phenomenon means that the most common word is twice as common as the second most common word, three times as common as the third most common word, four times as common as the fourth most common word, and so forth.
```

## A3. Analyzing word and document frequency: TF*IDF

###  --->>>  [looking at tokenized abstracts ]
The goal is to quantify what a document is about. What is the document about?

-   **term frequency (tf)** = how frequently a word occurs in a document... but there are words that occur many time and are not important
-   term's **inverse document frequency (idf)** = decreases the weight for commonly used words and increases the weight for words that are not used very much in a collection of documents.
-   **statistic tf-idf (= tf*-*idf)** = tan alternative to using stopwords is _the frequency of a term adjusted for how rarely it is used_. [It measures how important a word is to a document in a collection (or corpus) of documents, but it is still a rule-of-thumb or heuristic quantity] 

> The inverse document frequency for any given term is defined as:

$$idf(\text{term}) = \ln{\left(\frac{n_{\text{documents}}}{n_{\text{documents containing term}}}\right)}$$

```{r}
# PREP X TF-IDF
paint(abs_all)
skimr::n_unique(abs_all$title)
skimr::n_unique(abs_all$date_issued)

# Count words (with stopwords)
abs_words <-  abs_all %>% 
  unnest_tokens(output =  word,
                input = abstract,
                to_lower = T, # otherwise cannot match the stop_words  
                strip_punc = TRUE) %>%  #9769 = tot words
  # other important pre-process step
  mutate(title = factor(title, ordered = TRUE))  %>% 
  mutate(date_issued = factor(date_issued, ordered = TRUE)) %>% 
  # implicit group by ...
  group_by(date_issued) %>%
  count(word, sort = TRUE) %>% # 5860 # unique words
  ungroup() %>% 
  left_join(abs_all, by = "date_issued")

```


### modo I) {LinkedIn GUY} Create a Word Frequency Table
[tm way](https://www.linkedin.com/learning/processing-text-with-r-essential-training/improving-term-frequency-matrix?autoSkip=true&autoplay=true&contextUrn=urn%3Ali%3AlearningCollection%3A6932688789001674752&resume=false)
```{r, eval=FALSE}
abs_words_freq <-  abs_words %>% 
  anti_join(stop_words, by= "word" )  %>% 
  tidyr::pivot_wider(names_from = word, values_from = n, values_fill = 0)

# skim(abs_words_freq) 
 
#Generate the Document Term matrix
abs_words_freq_matrix <- as.matrix(abs_words_freq)

abs_words_freq_matrix[ , 'gender']

# .... uses {tm} 
```


### modo II) {Julia Silge and David Robinson} tidytext::bind_tf_idf 
[tidytext way](https://www.linkedin.com/learning/processing-text-with-r-essential-training/improving-term-frequency-matrix?autoSkip=true&autoplay=true&contextUrn=urn%3Ali%3AlearningCollection%3A6932688789001674752&resume=false)

The idea of `tf-idf` is to find the important words for the content of each document by _decreasing the weight for commonly used words and increasing the weight for words that are not used very much in a collection or corpus of documents_.
Calculating tf-idf attempts to find the words that are important (i.e., common) in a text, but not too common. Let’s do that now.

```{r prep_tf_idf}
paint(abs_words)
### ---  abstracts with totals 

# Calculate the total appearances of each words per doc
abs_total_words <- abs_words %>%
  dplyr::group_by(title) %>%
  dplyr::summarise(total = sum(n))  

# Join the total appearances of each words per doc
abs_words_T <- left_join(abs_words, abs_total_words) %>% 
  select(date_issued,title,abstract, word, n, total )
paint(abs_words_T)

# The usual suspects are here, “the”, “and”, “to”, and so forth. 
# ggplot(abs_words_T, aes(n/total, fill = title)) +
#   geom_histogram(show.legend = FALSE) +
#    xlim(NA, 0.01) +
#   facet_wrap(~title, ncol = 2, scales = "free_y")

```

>`tidytext::bind_tf_idf`: Calculate and bind the term frequency and inverse document frequency of a tidy text dataset, along with the product, tf-idf, to the dataset. Each of these values are added as columns. This function supports non-standard evaluation through the tidyeval framework.

```{r tf_idf}
abs_words_tf_idf <- abs_words_T %>%
  #bind_tf_idf(tbl, term, document, n)
  bind_tf_idf(      word, title, n) %>% # 5860
  # get rid of stopwords anyway &...
  anti_join(stop_words, by = "word") %>% # 3703
  # fileter most importantly weighted 
  # filter(tf_idf > 0.01 ) %>% # 2278
  arrange(date_issued, desc(tf_idf))

abs_words_tf_idf
paint(abs_words_tf_idf)
```

Notice that `idf` and thus `tf-idf` are zero for these extremely common words. These are all words that appear in all docs, so the `idf` term (which will then be the natural log of 1) is zero. 
> The *inverse document frequency* (and thus `tf-idf`) is very low (near zero) for words that occur in many of the documents in a collection; this is how this approach decreases the weight for common words. 

> The *inverse document frequency* will be a higher number for words that occur in fewer of the documents in the collection. 

Let’s look at terms with high tf-idf in WDRs.
```{r}
abs_words_tf_idf %>%
  select(-total) %>%
  arrange(desc(tf_idf))

# let's look specifically at "Gender Equality and Development" 
tf_idf_2012 <- abs_words_tf_idf %>%
  filter(date_issued == "2012") %>%
  select(-total) %>%
  arrange(desc(tf_idf))

# These words are, as measured by tf-idf, the most important to "Gender Equality and Development"  and most readers would likely agree.
tf_idf_2012[tf_idf_2012$word %in% c("gender", "equality", "development") ,]

# # A tibble: 3 × 7
#   date_issued title                              word            n     tf   idf tf_idf
#   <chr>       <chr>                              <chr>       <int>  <dbl> <dbl>  <dbl>
# 1 2012        " Gender Equality and Development" gender         13 0.0588  3.09  0.182
# 2 2012        " Gender Equality and Development" equality        6 0.0271  3.78  0.103
# 3 2012        " Gender Equality and Development" development    10 0.0452  0     0 
```

### ---  TF*IDF tables for selected WDRs (1992)
```{r tab_env}
skim(abs_words_tf_idf$tf_idf)

# SIMPLE TABLE WITH FILTER 
tf_idf_1992_2010  <-  abs_words_tf_idf %>% 
  dplyr::filter(date_issued %in%  c("1992", "2010")) %>% 
    dplyr::filter(n > 1) %>% 
  select(-abstract) %>%
  dplyr::arrange(date_issued, desc(tf_idf)) 

tf_idf_1992_2010 %>%
  DT::datatable(.,
                elementId = NULL,
                filter = 'top',  # Problem if there are empty cells !!!
                extensions = 'Buttons',
                options = list(columnDefs = list(list(targets = c(1, 2, 3), searchable = TRUE)),
                               pageLength = 50,
                               searchHighlight = TRUE,
                               dom = 'Bfrtip',
                               buttons = c('csv','excel','pdf'))) %>%
    # helper functions
        DT::formatRound(c('tf', 'idf', 'tf_idf'), 3) %>%
        DT::formatStyle('tf_idf', backgroundColor =  'yellow') %>%
        DT::formatStyle('tf_idf', color = styleInterval( 0.0114, c('black', 'red')) )

 
# tf_idf_1992_2010   %>% 
#   # alternative with filters
#   datatable(
#   tf_idf_1992_2010,
#   filter = 'top',
#   options = list(columnDefs = list(list(targets = c(1, 2, 3), searchable = TRUE)),
#     pageLength = 5
#   )
# )
```

```{r tab_know}
skim(abs_words_tf_idf$tf_idf)

# SIMPLE TABLE WITH FILTER 
tf_idf_1998_2021  <-  abs_words_tf_idf %>% 
  dplyr::filter(date_issued %in%  c("1998", "2021")) %>% 
    dplyr::filter(n > 1) %>% 
  select(-abstract) %>%
  dplyr::arrange(date_issued, desc(tf_idf)) 

tf_idf_1998_2021 %>%
  DT::datatable(.,
                elementId = NULL,
                filter = 'top',  # Problem if there are empty cells !!!
                extensions = 'Buttons',
                options = list(columnDefs = list(list(targets = c(1, 2, 3), searchable = TRUE)),
                               pageLength = 50,
                               searchHighlight = TRUE,
                               dom = 'Bfrtip',
                               buttons = c('csv','excel','pdf'))) %>%
    # helper functions
        DT::formatRound(c('tf', 'idf', 'tf_idf'), 3) %>%
        DT::formatStyle('tf_idf', backgroundColor =  'yellow') %>%
        DT::formatStyle('tf_idf', color = styleInterval( 0.0114, c('red', 'blue')) )

 
# tf_idf_1992_2010   %>% 
#   # alternative with filters
#   datatable(
#   tf_idf_1992_2010,
#   filter = 'top',
#   options = list(columnDefs = list(list(targets = c(1, 2, 3), searchable = TRUE)),
#     pageLength = 5
#   )
# )
```


### ---  Word  frequency plots
```{r}
# let's eliminate stopwords 
abs_words2 <-  anti_join(x = abs_words_T, y = stop_words, by= "word" )  %>% 
  filter(n > 1) %>% 
 select(date_issued, title, word, n, total)

abs_words2 
 paint(abs_words2)
# here there is one row for each word-WDR(abs) combination 
# `n` is the number of times that word is used in that book and 
# `total` is the total words in that book.
```

### ---  Word  frequency plots
 ***frequency*** = let's look at the distribution of `n/total` for each novel, the number of times a word appears in a novel divided by the total number of terms (words) in that novel. 

> I actually use `n` instead because I have only small numbers having used the abstracts alone

```{r hist_one}
one <- abs_words2 %>% 
  filter ( date_issued == "2021") %>% 
  mutate (freq = n/total)

# overlayed mess!
ggplot(data = one, 
       mapping = aes(x = n, fill = title)) + # y axis not needed ... R will count
  geom_histogram(binwidth = 1,
                 color = "white") +
  scale_y_continuous(breaks= pretty_breaks()) +
  xlim(0, 20) +
  labs(# title =  title, 
    x = "frequency",
    y = "N of words @ that frequency") + 
  guides( fill = "none")
```

```{r hist_mess, eval=FALSE}
# overlayed mess!
ggplot(abs_words2, aes(n, fill = title)) +
  geom_histogram(binwidth = 1,
                 color = "white") +
  scale_y_continuous(breaks= pretty_breaks()) +
  xlim(0, 20) +
  labs(#title = ~date_issued, 
    x = "frequency",
    y = "N of words @ that frequency") + 
  guides( fill = "none")
```

```{r hist_facet, eval=FALSE}
ggplot(abs_words2, aes(n, fill = title)) +
  geom_histogram(binwidth = 1,
                 color = "white") +
  scale_y_continuous(breaks= pretty_breaks()) +
  xlim(0, 20) +
  labs(#title = ~date_issued, 
    x = "frequency",
    y = "N of words @ that frequency") + 
  facet_wrap( ~date_issued ) + # , ncol = 2, scales = "free_y")
  guides( fill = "none") # way to turn legend off
```

#### --- Multiple plots of Word Freq with [purrr] 

+ [Jennifer Thompson, An Intro to the Magic of purrr](https://htmlpreview.github.io/?https://github.com/jenniferthompson/RLadiesIntroToPurrr/blob/master/purrr_intro.nb.html)

+ [automate ggplots while using variable labels as title and axis titles](https://stackoverflow.com/questions/59095237/automate-ggplots-while-using-variable-labels-as-title-and-axis-titles)

```{r hist_purrr}
# ---- NON Capisco 
# # Preferred approach
# histos <- abs_words2 %>%
#   group_by(title) %>%
#   nest() %>%
#   mutate(plot = map2(data, title, 
#                      ~ggplot(data = .x , aes(n/total, fill = title)) + 
#                       geom_histogram( show.legend = FALSE) + 
#                        xlim(NA, 0.05) +
#                        # ggtitle(.y) +
#                        ylab("Words Frequency") +
#                        xlab("Distribution per WDR title")))
# 
#  histos$plot[[1]]


# ---- Capisco 
# https://stackoverflow.com/questions/60671725/ggplot-add-title-based-on-the-variable-in-the-dataframe-used-for-plotting
list_plot <- abs_words2 %>%
  dplyr::group_split(date_issued) %>% # Split data frame by groups
  map(~ggplot(., 
              mapping = aes(x = n, fill = title)) +
        geom_histogram(binwidth = 1,
                       color = "white") +
        scale_y_continuous(breaks= pretty_breaks()) + # integer ticks {scales}
        xlim(0, 18) + # max of freq is 17
        labs(title = .$date_issued, 
             x = "frequency",
             y = "N of words @ that frequency")
  )

list_plot[[1]]
list_plot[[44]]
#grid.arrange(grobs = list_plot, ncol = 1)
```


#### -------------- STOP ------------------

### --->>> [looking at SUBJECTS ]



## A2.b Tokenization by n-gram - ITERATIVELY]

### -- using abstract? 
### -- using subjects? 

# ---


## A4. Relationships between words: n-grams and correlations

examples are examining which words tend to follow others immediately, or that tend to co-occur within the same documents. + `tidytext::token = "ngrams" argument` is a method tidytext offers for calculating and visualizing relationships between words in your text dataset. It tokenizes by pairs of adjacent words rather than by individual ones. We'll also introduce two new packages: `ggraph`, which extends `ggplot2` to construct network plots, and `widyr`, which calculates pairwise correlations and distances within a tidy data frame. Together these expand our toolbox for exploring text within the tidy data framework.

### --- Tokenizing by n-gram

```{r}
# break the text [e.g. 1 abstract] into individual tokens 
abs_1 <- as_tibble(wdr$abstract[1] )
abs_1_t <-  abs_1 %>% 
  # unnest (newColName, oldColName)
  unnest_tokens(word, value ) # --> tidytext


abs_2022_bigrams <- wdr %>% 
  dplyr::filter(id == 36883) %>% 
  select(abstract) %>% 
  as_tibble() %>% 
  unnest_tokens(., output = bigram, input = abstract, token = "ngrams", n=2 )  
# notice how these bigrams overlap
```

### --- Tokenizing by n-gram - ITERATIVELY

```{r}
# ---modo 1
# here I am basically adding rows wit repeated id attached 
abs_bigrams <- wdr %>% 
  #dplyr::filter(id == 36883) %>% 
  select(id, title, date_issued, abstract) %>% 
  as_tibble() %>% 
  unnest_tokens(., output = bigram, input = abstract, token = "ngrams", n=2 )  %>% 
  count(bigram, sort = TRUE) # 7348 obs 

# ----- what if I wanted to keep my 45 rows and add bigrams in a wide sense? 

# ---modo 2
 abs_prep <- wdr %>% 
  #dplyr::filter(id == 36883) %>% 
  select(id, title, date_issued, abstract) %>% 
  as_tibble() # 45 obs
 
myComplexFunction <- function(arg1, arg2, arg3, arg4){
  col <- arg1 * arg2 * arg3 * arg4
  x <- as.data.frame(col)
}

arg1 <- c(1, 3, 5, 7, 9)
arg2 <- c(2, 4, 6, 8, 10)
arg3 <- c(5, 10, 15, 20, 25)
arg4 <- c(10, 20, 30, 40, 50)

argList <- list(arg1, arg2, arg3, arg4)

# error Error in is_grouped_df(tbl) : argument "tbl" is missing, with no default
# abs_bigr_45 <- map_dfr(tbl = abs_prep,
#                        .x = abs_prep$abstract, .f = unnest_tokens(), output = bigram, input = abstract, token = "ngrams", n=2 )


```

### --- Counting and filtering n-grams

```{r}
# separate words 
abs_2022_bigrams_sep <- abs_2022_bigrams %>% 
  count(bigram, sort = TRUE) %>%
  # since many are bigram with a stopword
  separate(bigram, c("word1", "word2"), sep = " ") %>% 
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word)

# new bigram counts:
abs_2022_bigrams_sep_counts <- abs_2022_bigrams_sep %>% 
  count(word1, word2, sort = TRUE)

abs_2022_bigrams_sep_counts

 # OPPOSITE: reunite words 
abs_2022_bigrams_united <- abs_2022_bigrams_sep %>% 
  unite(bigram, word1, word2, sep = " ")
  
```

### --- Analyzing bigrams

In other analyses you may be interested in the most common trigrams, which are consecutive sequences of 3 words. We can find this by setting n = 3:

```{r}

```

### --- Trigrams

In other analyses you may be interested in the most common trigrams, which are consecutive sequences of 3 words. We can find this by setting n = 3:

```{r}

```

## A5. [Concordances: lag and lead words...]

As in [Ben Schmidt Chp 8.2.3](http://benschmidt.org/HDA/texts-as-data.html#concordances)

# -------------- TO do \| To Rethink ------------------

+ create my own stop_words list (which excludes also "date_issued", "report", etc)
+ leggere [@yusuf_development_2008]

### ----------------------------------------------------------------------------

# B. World Bank Projects & Operations:

# Connections

Following the example of [David Robinson on HN titles](http://varianceexplained.org/r/hn-trends/)

# Data sources:

1.  World Bank Projects & Operations: <https://datacatalog.worldbank.org/search/dataset/0037800> <https://datacatalog.worldbank.org/search/dataset/0037800/World-Bank-Projects---Operations>

-   Accessibility Classification: **public** under [Creative Commons Attribution 4.0](https://datacatalog.worldbank.org/public-licenses?fragment=cc)

2.  World Bank - World Development Reports

-   Accessibility Classification:

# Reference Tutorials

@robinson_1_2022 [Benjamin Soltoff: Computing 4 Social Sciences - API](https://cfss.uchicago.edu/syllabus/getting-data-from-the-web-api-access/) [Benjamin Soltoff: Computing 4 Social Sciences - text analysis](https://cfss.uchicago.edu/syllabus/text-analysis-fundamentals-and-sentiment-analysis/)

[Ben Schmidt Book Humanities Crurse](https://hdf.benschmidt.org/R/) [Ben Schmidt Book Humanities](http://benschmidt.org/HDA/texts-as-data.html)


