---
title: "Process and merge data"
author: "Luisa M. Mimmi"
date: "Last run: `r format(Sys.time(), '%F')`"
output: 
  html_document:
    code_folding: hide
editor_options: 
  chunk_output_type: console
bibliography: buzzwords.bib
---

<i class="fa fa-refresh" style="color: firebrick"></i> Work in progress

```{r setup_inp}
# Pckgs -------------------------------------
if (!require ("pacman")) (install.packages("pacman"))
p_load(tidyverse, 
       # ggplot2, for data visualisation.
       # dplyr, for data manipulation.
       # tidyr, for data tidying.
       # readr, for data import.
       # purrr, for functional programming.
       # tibble, for tibbles, a modern re-imagining of data frames.
       # stringr, for strings.
       # forcats, for factors.
       here,
       httr, jsonlite, XML, xml2, 
       oai, # R client to work with OAI-PMH 
       citr, 
       tidytext)

# Function(s) -------------------------------------
# source(here::here ("R", "WB-OKR_scrape-function.R") )

# Data -------------------------------------
# DATA https://datacatalog.worldbank.org/search/dataset/0037800 
# INSTRUCTIONS https://documents.worldbank.org/en/publication/documents-reports/api

```

# Data ingesting & cleaning

## 1. World Development Reports (WRDs) 
<!-- Done in "./_my_stuff/_01_data-overview.Rmd"  -->

```{r}
# WRD metadata taken with API get (issues) 
WDR <- readr::read_rds(here::here("data", "raw_data", "WDR.rds" )) %>% 
      # Extract only the portion of string AFTER the backslash {/}
  mutate(id =  as.numeric(stringr::str_extract(doc_mt_identifier_1, "[^/]+$")) ) %>% 
  dplyr::relocate(id, .before = doc_mt_identifier_1) %>% 
  mutate(url_keys = paste0("https://openknowledge.worldbank.org/handle/10986/", id , "?show=full") )%>% 
  dplyr::relocate(url_keys, .before =  doc_mt_identifier_1)

# WRD subject/date_issued taken by manual review 
WDR_subjects <- readxl::read_excel(here::here("data", "raw_data", "WDR_subjects.xlsx" )) %>% 
  drop_na(id)

# join
WDR_com <- left_join(WDR, WDR_subjects, by = "id") %>% 
  dplyr::relocate(date_issued, .before = id ) %>% 
  # drop useles clmns 
  dplyr::select(-doc_mt_identifier_1, -doc_mt_identifier_2, -doc_mt_date, 
                -doc_mt_subject, -doc_mt_creator, -doc_mt_set_spec) %>% 
  dplyr::relocate(url_keys, .after = subj_216 ) %>% 
  dplyr::rename(abstract = doc_mt_description) %>% 
  # correct titles -> portion after {:}
  dplyr::mutate(., title = str_extract(doc_mt_title,"[^:]+$")) %>% 
  dplyr::relocate(title, .after = id)  %>% 
  dplyr::rename(title_miss = doc_mt_title) %>% 
  dplyr::mutate(title_miss = case_when(
      str_starts(title, "World Development Report") ~ "Y",
      TRUE ~ NA_character_) 
  ) %>% 
  dplyr::mutate(subject_mess = if_else(!is.na(subj_136), 
                                       "Y", 
                                       NA_character_))  %>% 
  dplyr::relocate(subject_mess, .after = title_miss)  
 
#convert all columns that start with "subj_" to lowercase
# WDR_com[7:223] <- sapply(WDR_com[7:223], function(x) tolower(x))
WDR_com[, grep("^subj_", names(WDR_com))] <- sapply(WDR_com[, grep("^subj_", names(WDR_com))], function(x) tolower(x))

# combine all `subj_...` vars into a vector separated by comma
col_subj <- names(WDR_com[, grep("^subj_", names(WDR_com))] )

WDR_com <- WDR_com %>% tidyr::unite(
  col = "subj_all", 
  subj_1:subj_216, 
  sep = ",",
  remove = FALSE)
```


### Wrangling the text

Following: [@kaye_ella_2019; @robinson_words_2017; @robinson_1_2022]

### Tokenization

Some definitions:
-   A **token** ~ is a meaningful unit of text, such as a *word*, that we are interested in using for analysis
-   *tokenization* ~ splitting text into tokens
-   **bigrams**/**n-gram** ~ sentence, or paragraph
-   ***tidy text*** ~ "a table with one-token-per-row"
-   ***corpus*** ~  objects that typically contain raw strings annotated with additional metadata and details
-   ***Document-term matrix*** ~ sparse matrix describing a collection (i.e., a corpus) of documents with one row for each document and one column for each term.  (`WDR_com`)
 
```{r}
# printing out data as tibble 
sbj_all_4 <- c(WDR_com$subj_1[4],WDR_com$subj_2[4] ,WDR_com$subj_3[4] ,WDR_com$subj_4[4] )

library(dplyr)
sbj_all_4_df <- tibble(line = 1:4, sbj_all_4 = sbj_all_4)


# break the text [e.g. 1 abstract] into individual tokens 
abs_1 <- as_tibble(WDR_com$abstract[1] )
abs_1_t <-  abs_1 %>% 
  # unnest (newColName, oldColName)
  unnest_tokens(word, value ) # --> tidytext
```
 
 
# -------------- REVIEW ------------------ #

## 2. World Bank Projects & Operations:


# Connections

Following the example of [David Robinson on HN titles](http://varianceexplained.org/r/hn-trends/)

# Data sources:

1.  World Bank Projects & Operations: <https://datacatalog.worldbank.org/search/dataset/0037800> <https://datacatalog.worldbank.org/search/dataset/0037800/World-Bank-Projects---Operations>

-   Accessibility Classification: **public** under [Creative Commons Attribution 4.0](https://datacatalog.worldbank.org/public-licenses?fragment=cc)

2.  World Bank - World Development Reports

-   Accessibility Classification:

# Tutorials

[Benjamin Soltoff: Computing 4 Social Sciences - API](https://cfss.uchicago.edu/syllabus/getting-data-from-the-web-api-access/) [Benjamin Soltoff: Computing 4 Social Sciences - text analysis](https://cfss.uchicago.edu/syllabus/text-analysis-fundamentals-and-sentiment-analysis/)
