---
title: "Process and merge data"
author: "Luisa M. Mimmi"
date: "Last run: `r format(Sys.time(), '%F')`"
output: 
  html_document:
    code_folding: hide
editor_options: 
  chunk_output_type: console
bibliography: buzzwords.bib
---

<i class="fa fa-refresh" style="color: firebrick"></i>  Work in progress
 


```{r setup_inp}
# Pckgs -------------------------------------
if (!require ("pacman")) (install.packages("pacman"))
p_load(tidyverse, 
       # ggplot2, for data visualisation.
       # dplyr, for data manipulation.
       # tidyr, for data tidying.
       # readr, for data import.
       # purrr, for functional programming.
       # tibble, for tibbles, a modern re-imagining of data frames.
       # stringr, for strings.
       # forcats, for factors.
       here,
       httr, jsonlite, XML, xml2, 
       oai, # R client to work with OAI-PMH 
       citr)

# Function(s) -------------------------------------
# source(here::here ("R", "WB-OKR_scrape-function.R") )

# Data -------------------------------------
# DATA https://datacatalog.worldbank.org/search/dataset/0037800 
# INSTRUCTIONS https://documents.worldbank.org/en/publication/documents-reports/api

```
  
# Data ingest
<!-- Done in "./_my_stuff/_01_data-overview.Rmd"  -->

```{r}
# WRD metadata taken with API get (issues) 
WDR <- readr::read_rds(here::here("data", "raw_data", "WDR.rds" ))

# WRD subject/date_issued taken by manual review 
WDR_subjects <- readxl::read_excel(here::here("data", "raw_data", "WDR_subjects.xlsx" )) 
WDR_subjects <- WDR_subjects[1:45,]

```


### Parsing xml ONE INDICATOR AT A TIME (header)

 
# REVIEW --- 

## 1. World Bank Projects & Operations:

 

# Wrangling the text 
Following: [@kaye_ella_2019; @robinson_words_2017]


## Tokenization 

* A _token_ is a meaningful unit of text, such as a word, that we are interested in using for analysis
* bigrams 

# connections 

Following the example of [David Robinson on HN titles](http://varianceexplained.org/r/hn-trends/)
  
# Data sources: 

1. World Bank Projects & Operations: https://datacatalog.worldbank.org/search/dataset/0037800 
 https://datacatalog.worldbank.org/search/dataset/0037800/World-Bank-Projects---Operations

  + Accessibility Classification: **public** under [Creative Commons Attribution 4.0](https://datacatalog.worldbank.org/public-licenses?fragment=cc)

2. World Bank - World Development Reports

  + Accessibility Classification:

# Tutorials 
[Benjamin Soltoff: Computing 4 Social Sciences - API](https://cfss.uchicago.edu/syllabus/getting-data-from-the-web-api-access/)
[Benjamin Soltoff: Computing 4 Social Sciences - text analysis](https://cfss.uchicago.edu/syllabus/text-analysis-fundamentals-and-sentiment-analysis/)

