---
title: "Process and merge data"
author: "Luisa M. Mimmi"
date: "Last run: `r format(Sys.time(), '%F')`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_depth: 3
editor_options: 
  chunk_output_type: console
bibliography: slogan.bib
---

<i class="fa fa-refresh" style="color: firebrick"></i> Work in progress

#### setup
```{r setup_inp, echo=FALSE}
# Pckgs -------------------------------------
if (!require ("pacman")) (install.packages("pacman"))
#p_install_gh("luisDVA/annotater")
#p_install_gh("HumanitiesDataAnalysis/hathidy")
# devtools::install_github("HumanitiesDataAnalysis/HumanitiesDataAnalysis") 
pacman::p_load(here,
       annotater,
       tidyverse, 
       skimr,
       scales,
       httr, 
       citr,
       paint, 
       DT, # an R interface to the JavaScript library DataTables
       # jsonlite, 
       # XML, 
       # xml2, 
       # oai, # R client to work with OAI-PMH 
       splitstackshape #Stack and Reshape Datasets After Splitting Concatenated Values
       )

pacman::p_load(
  tm, # Text Mining Package
  tidytext, # Text Mining using 'dplyr', 'ggplot2', and Other Tidy Tools
  widyr, # pairwise ????  
  ggraph,
  igraph,
  sjmisc, # Data and Variable Transformation Functions
  ggraph, # An Implementation of Grammar of Graphics for Graphs and Networks
  widyr, # Widen, Process, then Re-Tidy Data
  SnowballC, # Snowball Stemmers Based on the C 'libstemmer' UTF-8 Library
  HumanitiesDataAnalysis, # Data and Code for Teaching Humanities Data Analysis
  sentencepiece # Text Tokenization using Byte Pair Encoding and Unigram Modelling
)

# Function(s) -------------------------------------

# Data -------------------------------------
 
```

#### -- Set stopwords [more...]

```{r stop}
# --- set up stop words
stop_words <- as_tibble(stop_words) %>% # in the tidytext dataset 
  add_row(word = "WDR", lexicon = NA_character_) %>%
  add_row(word = "world", lexicon = NA_character_) %>%
  add_row(word = "report", lexicon = NA_character_) %>%
  add_row(word = "development", lexicon = NA_character_) %>%
  add_row(word = "1978", lexicon = NA_character_) %>%
  add_row(word = "1979", lexicon = NA_character_) %>%
  add_row(word = "1980", lexicon = NA_character_) %>%
  add_row(word = "1981", lexicon = NA_character_) %>%
  add_row(word = "1982", lexicon = NA_character_) %>%
  add_row(word = "1983", lexicon = NA_character_) %>%
  add_row(word = "1984", lexicon = NA_character_) %>%
  add_row(word = "1985", lexicon = NA_character_) %>%
  add_row(word = "1986", lexicon = NA_character_) %>%
  add_row(word = "1987", lexicon = NA_character_) %>%
  add_row(word = "1988", lexicon = NA_character_) %>%
  add_row(word = "1989", lexicon = NA_character_) %>%
  add_row(word = "1990", lexicon = NA_character_) %>%
  add_row(word = "1991", lexicon = NA_character_) %>%
  add_row(word = "1992", lexicon = NA_character_) %>%
  add_row(word = "1993", lexicon = NA_character_) %>%
  add_row(word = "1994", lexicon = NA_character_) %>%
  add_row(word = "1995", lexicon = NA_character_) %>%
  add_row(word = "1996", lexicon = NA_character_) %>%
  add_row(word = "1997", lexicon = NA_character_) %>%
  add_row(word = "1998", lexicon = NA_character_) %>%
  add_row(word = "1999", lexicon = NA_character_) %>%
  add_row(word = "2000", lexicon = NA_character_) %>%
  add_row(word = "2001", lexicon = NA_character_) %>%
  add_row(word = "2002", lexicon = NA_character_) %>%
  add_row(word = "2003", lexicon = NA_character_) %>%
  add_row(word = "2004", lexicon = NA_character_) %>%
  add_row(word = "2005", lexicon = NA_character_) %>%
  add_row(word = "2006", lexicon = NA_character_) %>%
  add_row(word = "2007", lexicon = NA_character_) %>%
  add_row(word = "2008", lexicon = NA_character_) %>%
  add_row(word = "2009", lexicon = NA_character_) %>%
  add_row(word = "2010", lexicon = NA_character_) %>%
  add_row(word = "2011", lexicon = NA_character_) %>%
  add_row(word = "2012", lexicon = NA_character_) %>%
  add_row(word = "2013", lexicon = NA_character_) %>%
  add_row(word = "2014", lexicon = NA_character_) %>%
  add_row(word = "2015", lexicon = NA_character_) %>%
  add_row(word = "2016", lexicon = NA_character_) %>%
  add_row(word = "2017", lexicon = NA_character_) %>%
  add_row(word = "2018", lexicon = NA_character_) %>%
  add_row(word = "2019", lexicon = NA_character_) %>%
  add_row(word = "2020", lexicon = NA_character_) %>%
  add_row(word = "2021", lexicon = NA_character_) %>%
  add_row(word = "2022", lexicon = NA_character_) %>% 
  filter (word != "changes")

# --- set up stop words stemmed
stop_words_stem <- stop_words  %>% 
mutate (word = SnowballC::wordStem(word ))
```



# B. World Bank Projects & Operations:

# Connections

Following the example of [David Robinson on HN titles](http://varianceexplained.org/r/hn-trends/)

# Data sources:

1.  World Bank Projects & Operations: <https://datacatalog.worldbank.org/search/dataset/0037800> <https://datacatalog.worldbank.org/search/dataset/0037800/World-Bank-Projects---Operations>

-   Accessibility Classification: **public** under [Creative Commons Attribution 4.0](https://datacatalog.worldbank.org/public-licenses?fragment=cc)

2.  World Bank - World Development Reports

-   Accessibility Classification:

# Reference Tutorials

@robinson_1_2022 [Benjamin Soltoff: Computing 4 Social Sciences - API](https://cfss.uchicago.edu/syllabus/getting-data-from-the-web-api-access/) [Benjamin Soltoff: Computing 4 Social Sciences - text analysis](https://cfss.uchicago.edu/syllabus/text-analysis-fundamentals-and-sentiment-analysis/)

[Ben Schmidt Book Humanities Crurse](https://hdf.benschmidt.org/R/) [Ben Schmidt Book Humanities](http://benschmidt.org/HDA/texts-as-data.html)

[tidyTuesday cast on tidytext](https://github.com/dgrtwo/data-screencasts/tree/master/screencast-annotations)

  1. ✔️ [MEDIUM articles: common words, pairwise correlations - 2018-12-04](https://www.youtube.com/watch?v=C69QyycHsgE)
  2. [TidyTuesday Tweets -  2019-01-07](https://www.youtube.com/watch?v=KE9ItC3doEU)
  3. [Wine Ratings - 2019-05-31](https://www.youtube.com/watch?v=AQzZNIyjyWM) Lasso regression | sentiment lexicon,
  4. [Simpsons Guest Stars 	2019-08-30](https://www.youtube.com/watch?v=EYuuAGDeGrQ) geom_histogram
  5. [Horror Movies 	2019-10-22](https://www.youtube.com/watch?v=yFRSTlk3kRQ) explaining glmnet package | Lasso regression
  6. [The Office 	2020-03-16](https://www.youtube.com/watch?v=_IvAubTDQME) geom_text_repel from ggrepel | glmnet package to run a cross-validated LASSO regression
  7. [Animal Crossing 	2020-05-05](https://www.youtube.com/watch?v=Xt7ACiedRRI) Using geom_line and geom_point to graph ratings over time | geom_text to visualize what words are associated with positive/negative reviews |topic modelling
